Adam Prado
Pace U. Data Science
4/7/24

Chapter 7, Applied #6, R

---
title: "Chapter 7 Applied 6"
output: html_notebook
---

6)
Download the Wage dataset from the ISLR2 library,

```{r}
library(ISLR2)
attach(Wage)
head(Wage)
dim(Wage)
```

a) 
This is just checking the coefficient for polynomial models up to degree of 5.
```{r}
set.seed(1)
for(d in 1:5){
  fit <- lm (wage ~ poly(age,d), data = Wage)
  print(coef(summary(fit)))
}
```

a) This splits the Wage dataset in halves into a training and validation set
```{r}

set.seed(1)
train <- sample(1:3000,1500)
train_data <- Wage[train, ]
validation_data <- Wage[-train,]
dim(train_data)
dim(validation_data)
```
a) This goes through degrees of the polynomial model from 1 up to 10 using a for loop. It fits a model with each degree using the training data.  It then uses that model to to find the error when compared to the validation data.  It then prints out the degree and the mean error for each model.
```{r}

for(d in 1:10){
  fit <- lm (wage ~ poly(age,d), data = train_data)
  probs <- predict(fit, newdata = validation_data, type = "response")
  ME <- mean((wage - predict(fit, Wage))[-train]^2)
  print(paste(d, ":" ,ME ))
}  
```
a) The polynomial degree with the lowest mean error when tested against the validation set was a degree of 4, the 1620 error was the lowest of all tests.


a) This checks the same models using ANOVA
```{r}
fit.1 <-lm(wage ~ age, data = Wage)
fit.2 <-lm(wage ~ poly(age,2), data = Wage)
fit.3 <-lm(wage ~ poly(age,3), data = Wage)
fit.4 <-lm(wage ~ poly(age,4), data = Wage)
fit.5 <-lm(wage ~ poly(age,5), data = Wage)
fit.6 <-lm(wage ~ poly(age,6), data = Wage)
fit.7 <-lm(wage ~ poly(age,7), data = Wage)
fit.8 <-lm(wage ~ poly(age,8), data = Wage)
fit.9 <-lm(wage ~ poly(age,9), data = Wage)
anova(fit.1, fit.2, fit.3, fit.4, fit.5, fit.6, fit.7, fit.8, fit.9 )
```
a) The results are consistent with the previous test.  Degree 3 or 4 are the best choices for the models. 

Graph of the 4th degree polynomial
```{r}
fit <- lm (wage ~ poly(age,4), data = train_data)
agelims <-range(age)
age.grid <- seq(from= agelims[1], to=agelims[2])
preds <- predict(fit, newdata = list(age=age.grid), se = TRUE)
se.bands <-cbind(preds$fit + 2 * preds$se.fit, preds$fit - 2 * preds$se.fit)
par(mfrow = c(1,1), mar=c(4.5, 4.5, 1, 1) , oma = c(0,0,4,0))
plot (age, wage, xlim=agelims, cex=0.5, col="darkgrey")
title("Degree 4 Polynomial", outer=T)
lines(age.grid, preds$fit, lwd=2, col="blue")
matlines(age.grid, se.bands, lwd =1, col ="blue", lty=3)
```

b) This is an example of a 4 cut step function.  It uses cross validation to find the mean error when compared to the validation data set.
```{r}
c <- 4
table(cut(age,c))
fit <- lm(wage~ cut(age,c), train_data)
coef(summary(fit))
probs <- predict(fit, newdata = validation_data, type = "response")
ME <- mean((wage - predict(fit, Wage))[-train]^2)
print(paste(c, ":" ,ME ))  
  
```

b) This loops through number of cuts from 2 to 30, creates a step function for each, and then finds the mean error.  It prints each number of cuts and the corresponding MSE 
```{r}
for(c in 2:30){
  table(cut(age,c))
  fit <- lm(wage~ cut(age,c), train_data)
  coef(summary(fit))
  probs <- predict(fit, newdata = validation_data, type = "response")
  ME <- mean((wage - predict(fit, Wage))[-train]^2)
  print(paste(c, ":" ,ME ))  
 
}

```

b) lowest error at cut = 10 with a MSE of 1629.1

This then fits the step function model with 10 cuts and prints the coefficients of the model. 
```{r}
fit <- lm(wage~ cut(age,10), data=Wage)
coef(summary(fit))

```

```{r}

agelims <-range(age)
age.grid <- seq(from= agelims[1], to=agelims[2])
preds <- predict(fit, newdata = list(age=age.grid), se = TRUE)
se.bands <-cbind(preds$fit + 2 * preds$se.fit, preds$fit - 2 * preds$se.fit)
par(mfrow = c(1,1), mar=c(2, 2, 1, 1) , oma = c(0,0,4,0))
plot (age, wage, xlim=agelims, cex=0.5, col="darkgrey")
title("Step Function with 10 cuts", outer=T)
lines(age.grid, preds$fit, lwd=2, col="red")
matlines(age.grid, se.bands, lwd =1, col ="blue", lty=3)
```
b) Above is the plot of the step function with 10 cuts. 
